Field-informed Reinforcement Learning for Learning Large Scale Collective Tasks

### 2023-05-10 paper.tex
chore: remove different variants
chore: update the description of dqn with gnn

  \includegraphics[width=\textwidth]{images/architecture.pdf}

@@

In our case, we leveraged the property of \ac{GNN}s to have a dual interpretation, i.e., to function globally over the entire graph and locally only over the neighborhood. 
%
Therefore, we relied on the value-based approach \ac{DQN}~\cite{mnih2015playing} \ga{to introduce in many rl??} with two major modifications: 
\begin{enumerate}
  \item experience replay stores experiences in the form of \emph{graphs},
  \item the neural network used to compute the Q function is based on a \ac{GNN} with an \ac{MLP} downstream.
\end{enumerate}
The first point is a natural extension due to the fact that we work on graphs rather than simple values. 
%
For the second point,
 the use of \acp{GNN} allows us to define policies on a variable neighborhood  that is essential in such systems as this can vary based on the neighborhood policy,
 and it is known that \acp{GNN} have a certain ability to generalize to new structures and scale with different nodes. 
% 
Additionally, 
 using the overall graph compared to local experiences makes learning more stable as it reduces the non-stationarity of the environment perceived by each node.
% 
 During learning, it is as if we have a \emph{single} agent collectively choosing which action to take (i.e., resulting action graph). 
%
\ga{todo: describe the algorithm with the modifications}
\begin{algorithm}
  \KwIn{Environment $\mathcal{E}$, graph replay buffer $\mathcal{D}$, target network $\theta^{-}$, current network $\theta$, exploration strategy $\epsilon$}
  \KwOut{Trained DQN model $\theta$}

@@

  \While{not done}{
  Observe current state $s$;

  With probability $\epsilon$ select a random action $a$, otherwise select $a = \arg\max_{a'} Q(s,a';\theta)$;

  Execute action $a$ in the environment $\mathcal{E}$ and observe the reward $r$ and the next state $s'$;

  Store transition $(s,a,r,s')$ in $\mathcal{D}$;

  Sample a batch of transitions $(s_i,a_i,r_i,s'i)$ from $\mathcal{D}$;

  Compute the target Q-value for each transition: $y_i = r_i + \gamma \max{a'} Q(s'i,a';\theta^{-})$;

  Update the current network weights using gradient descent: $\theta \leftarrow \theta - \alpha \nabla{\theta} \frac{1}{N} \sum_i (y_i - Q(s_i,a_i;\theta))^2$;

  Every $C$ steps, update the target network weights: $\theta^{-} \leftarrow \theta$
  ;
  }
  \caption{Deep Q-Network (DQN)}
  \end{algorithm}

@@

\paragraph*{Variant A}
\paragraph*{Variant B}
\paragraph*{Variant C}

>>>>

  \includegraphics[width=\textwidth]{imgs/architecture.pdf}

@@

In our case, we leveraged the property of \ac{GNN}s to have a dual interpretation, i.e., to function globally over the entire graph and locally only over the neighborhood. 
%
Therefore, we relied on the value-based approach \ac{DQN}~\cite{mnih2015playing} \ga{to introduce in many rl??} with two major modifications: 
\begin{enumerate}
  \item experience replay stores experiences in the form of \emph{graphs},
  \item the neural network used to compute the Q function is based on a \ac{GNN} with an \ac{MLP} downstream.
\end{enumerate}
The first point is a natural extension due to the fact that we work on graphs rather than simple values. 
%
For the second point,
 the use of \acp{GNN} allows us to define policies on a variable neighborhood  that is essential in such systems as this can vary based on the neighborhood policy,
 and it is known that \acp{GNN} have a certain ability to generalize to new structures and scale with different nodes. 
% 
Additionally, 
 using the overall graph compared to local experiences makes learning more stable as it reduces the non-stationarity of the environment perceived by each node.
% 
 During learning, it is as if we have a \emph{single} agent collectively choosing which action to take (i.e., resulting action graph). 
%
\ga{todo: describe the algorithm with the modifications}
\begin{algorithm}
  \KwIn{Environment $\mathbb{E}$, graph replay buffer $\mathcal{D}$, target network $\theta^{-}$, current network $\theta$, exploration strategy $\epsilon$}
  \KwOut{Trained DQN model $\theta$}

@@

  \While{not done}{
  Observe current graph observations $G_o$;

  With probability $\epsilon$ select a random graph action $G_a$, 
  otherwise select $G_a =$ 
  $\{ v \in V, argmax_{a'v} Q(G_o, \theta)[v, a'v]\ ,  E \in G_o \}$;

  Execute the collective action $G_a$ in the environment $\mathbb{E}$ and observe a graph-level reward $G_r$ and the next state $G_o'$;

  Store transition $(G_o,G_a,G_r, G_o')$ in $\mathcal{D}$;

  Sample a batch of graph transitions $(G^i_o,G^i_a,G^i_r,G^i_o)$ 
   from $\mathcal{D}$ and merge them in $(G_o,G_a,G_r,G_o)$;

  Compute the target Q-value for each node in the graph: 
  $y = G_r + \gamma * \max{a'} Q(s'i,a';\theta^{-})$;

  Compute the current expetected value for each node in the graph:
  $y^* = Q(s'i,a';\theta)$;

  Update the current network weights using gradient descent: $\theta \leftarrow \theta - \alpha \nabla{\theta} \frac{1}{|G_o|} (y - y^*)^2$;

  Every $C$ steps, update the target network weights: $\theta^{-} \leftarrow \theta$
  ;
  }
  \caption{Deep Q-Network (DQN) with GNN and Graph Replay Buffer}
  \end{algorithm}

@@

%\paragraph*{Variant A}
%\paragraph*{Variant B}
%\paragraph*{Variant C}
